---
title: "ANOVA for Journalists: Analyzing College Admissions Test Scores"
date: 11-3-2025
---

Luke Tabacco

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction

In this exercise, we'll explore a common journalism question: *Are there significant differences in standardized test scores across different types of colleges?*

We'll use ANOVA (Analysis of Variance) to determine if the observed differences in SAT scores across three college types (Public, Private, and Community College) represent real patterns or just random variation.

## Task 1: Load Libraries and Create Sample Data

```{r}
library(tidyverse)
```

```{r create-data}
# Set seed for reproducibility
set.seed(123)

# Create simulated SAT scores for admitted students across three college types
# Scale is 400-1600 for total SAT
# rnorm generates random numbers, in this case, 100 numbers
public_college <- rnorm(100, mean = 1350, sd = 120)
private_college <- rnorm(100, mean = 1450, sd = 100)
community_college <- rnorm(100, mean = 1050, sd = 150)

# Combine into a single dataframe
admissions_data <- data.frame(
  college_type = factor(rep(c("Public", "Private", "Community"), each = 100)),
  sat_score = c(public_college, private_college, community_college)
)

# Ensure SAT scores are within realistic range (400-1600)
admissions_data$sat_score <- pmin(pmax(admissions_data$sat_score, 400), 1600)

# Quick look at the data
head(admissions_data)
```

## Task 2: Calculate Basic Statistics

```{r summary-stats}
# Summary statistics by college type
admission_summary <- admissions_data |>
  group_by(college_type) |>
  summarize(
    mean_score = mean(sat_score),
    median_score = median(sat_score),
    sd_score = sd(sat_score),
    min_score = min(sat_score),
    max_score = max(sat_score)
  )

print(admission_summary)
```

Context: Fall 2024 midpoint SAT for University of Maryland was 1,465.


**Task 3** Describe the data

Write a couple of sentences describing the differences in SAT scores between types of colleges, using the descriptive statistics we just generated.


Looking at the statistics we just generated from SAT scores between different colleges, it cand definitely be seen that there is a difference between scores coming from the three different types of colleges. Starting with community college, they had a mean score of 1068.070 and a median score of 1055.387. Looking now at public schools, they had a mean score of 1360.621 with a median score of 1357.411. Finally, private universities had a mean score of 1435.026 with a median of 1427.417. 


## Task 4: Visualize the Data

Let's create visualizations to see the patterns:

```{r visualize-distributions}
# Create density plot showing overlapping distributions
ggplot(admissions_data, aes(x = sat_score, fill = college_type)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "SAT Score Distributions by College Type",
    subtitle = "Are these differences statistically significant?",
    x = "SAT Score (400-1600 scale)",
    y = "Density",
    fill = "College Type"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

```{r boxplot}
# Create boxplot for another view
ggplot(admissions_data, aes(x = college_type, y = sat_score, fill = college_type)) +
  geom_boxplot() +
  labs(
    title = "SAT Scores by College Type",
    x = "College Type",
    y = "SAT Score"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

## Task 5: Perform ANOVA Test

```{r anova-test}
# Run one-way ANOVA
anova_result <- aov(sat_score ~ college_type, data = admissions_data)
summary(anova_result)
```

The F-value is huge (284.3). The F-value measures the ratio of between-group variance to within-group variance, and this high value suggests the differences between college types far outweigh the variation within each type. The p-value is very low, indicating the differences between college types are statistically significant at any reasonable significance level. The three asterisks (***) highlight that this is significant at the 0.001 level. So we can say that with confidence that there are real differences in SAT scores between the three types of colleges.

How do we know the F-value is huge?
We can calculate the critical value for an F statistic by using the qf() function in R

```{r}
# Your ANOVA has:
# df1 (between groups) = 2
# df2 (within groups/residuals) = 297
# Common alpha levels: 0.05, 0.01, 0.001

# Critical values at different significance levels:
critical_f_05 <- qf(0.95, df1 = 2, df2 = 297)
critical_f_01 <- qf(0.99, df1 = 2, df2 = 297)
critical_f_001 <- qf(0.999, df1 = 2, df2 = 297)

cat("Here are the Critical F-values:\n")
cat("alpha of α = 0.05: ", round(critical_f_05, 3), "\n")
cat("alpha of α = 0.01: ", round(critical_f_01, 3), "\n")
cat("alpha ofα = 0.001:", round(critical_f_001, 3), "\n")
cat("The F-statistic is: 284.3\n")
cat("Ratio to most extreme critical value (α=0.001) to the F statistic:", round(284.3/critical_f_001, 1), "x\n")
```


## Task 5: Compare the differences between specific pairs

```{r post-hoc}
# Tukey's HSD test to see which specific pairs of college types differ
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)
```

The Tukey multiple comparisons test shows highly significant differences in SAT scores between all college types:

* Private vs. Community: Private colleges have SAT scores that are 366.96 points higher than community colleges on average (p < 0.0001)
* Public vs. Community: Public colleges have SAT scores that are 292.55 points higher than community colleges on average (p < 0.0001)
* Public vs. Private: Public colleges have SAT scores that are 74.40 points lower than private colleges on average (p = 0.000021)

All comparisons show extremely small p-values, indicating these differences are statistically significant at any conventional significance level. The 95% confidence intervals (shown in the "lwr" and "upr" columns) do not include zero for any comparison, further confirming these differences are statistically significant.

## Task 6: Report the Findings

Based on our analysis, we can now answer:

1. Is there a significant difference in SAT scores across college types?

ANSWER: Yes, there is a significant difference in SAT scores across all college types. Looking at the most recent test we conducted, there is a 366.96 point difference between private universities and community college, there is a 292.55 point difference between public schools and community college. Finally, there is a 74.4 point difference between public schools and private schools. Every one of these tests had a very low p-value, further indicated the authenticity of theses tests. 

2. Which specific college types differ from each other?

ANSWER: All three specific college types differ from each other. Some, however, differ more than others. Private schools and community college have the biggest difference of 366.96 points. Public schools and community college have the second greatest difference of 292.55 points. And finally, while not as large as the other two, private schools and public schools have a difference of 74.4 points.

3. How large are these differences?

ANSWER: Some of these differences are very large SAT score wise, one however was not too large. The difference between community college with public and private schools are very large. But, the difference between public and private schools were no where near as large. 

Answer those questions, providing evidence for your answers, and then write a headline and a lead paragraph reporting these findings using active voice and focusing on the differences across types.

HEADLINE AND LEAD:

Private Schools hold the greatest level of standardized test scores, with community colleges holding the lowest. 

Tests show that the value and average of SAT scores are scattered for different types of colleges. Private schools have the greatest average of SAT scores, followed by public schools which are less than 100 points behind, followed by community colleges. 

