---
title: "HW7_MandMs"

Luke Tabacco

---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv

```{r}
mm_dataset <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset

```{r}
sample_5 <- mm_dataset |> 
  sample_n(5)

```

### Task 4: Calculate the mean age and standard deviation of the sample, using either the blue or white number of M&Ms

```{r}
sample_5 |> summarize(mean=mean(red), sd = sd(red))
```

### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
mm_dataset |> summarize(mean = mean(red), sd = sd(red))

```

### Task 6: Compare the sample and population statistics

Compare the two sets of means and standard deviations. Are they similar?

A: Yes, the two sets of means and standard deviations are very similar. They're actually almost identical. The mean of the sample data was 72.6, while the mean of the full dataset was 72.98. Next, the standard deviation of the sample data was 16.27, while the standard deviation of the fully M&M dataset was 16.46. 

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

A: From the activity that we performed in class, to the homework that I just completed, i am learning that random sampling of data can actually be very useful, and very close to accurate. During the in class assignment, my sample data was also pretty close to the full population data. Now, looking at these calculations, the data is less that 1 unit off for both the mean and the standard deviation. Now, it is hard to tell if something like sampling will always be accurate because at the end of the day it is random. But, it does make you think that the act of random sampling in a dataset such as this one can be very useful and pretty accurate. 


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
