---
title: "Final Project - Part 1"
output: html_notebook
---

Luke Tabacco

```{r}
library(tidyverse)
```
```{r}
tuition_statistics <- read_csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv") 
```
Direct link to the data descriptions:https://cran.r-project.org/web/packages/collegeScorecard/collegeScorecard.pdf

Direct source link: https://collegescorecard.ed.gov/

What does it contain and what time period does it cover?

ANSWER: The data contains 255 different universities, providing various years of data throughout. Each university is provided with data such as the state in which the university resides. Furthermore, each data set for the universities provide the in-state tuition fee, out-of-state tuition fee, and the total undergraduate enrollment for the school years 2021-2022, 2022-2023, and 2023-2024. 

Which columns will be most important for your analysis?

ANSWER: The columns that are going to be most important toward my analysis are honestly all of them. I am going to be deep diving into how the differences of in-state and out-of-state tuition fees affect the total enrollment of different universities. So, all of the data columns that have been provided are going to be an important factor in terms of my analysis. The only one that I probably won't dive too deep are the states in which the universities reside. 

What's missing that you might need or want?

ANSWER: Honestly, something that is missing that I might want to include is data about the specific in and out of state tuition enrollments. More specifically, I would really like to have data about who in terms of the total enrollment are in-state, vs who is out-of-state. 

Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values. 

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2023_24_undergrad_enrollment`, na.rm = TRUE),
          sd_stats = sd(`2023_24_undergrad_enrollment`, na.rm = TRUE),
          min_stats = min(`2023_24_undergrad_enrollment`, na.rm = TRUE),
          max_stats = max(`2023_24_undergrad_enrollment`, na.rm = TRUE)
              )
```

-> Why I calculated this and its importance at the bottom after all the calculations are done!

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2023_24_tuitionfee_in`, na.rm = TRUE),
          sd_stats = sd(`2023_24_tuitionfee_in`, na.rm = TRUE),
          min_stats = min(`2023_24_tuitionfee_in`, na.rm = TRUE),
          max_stats = max(`2023_24_tuitionfee_in`, na.rm = TRUE)
              )
```
-> Why I calculated this and its importance at the bottom after all the calculations are done!

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2023_24_tuitionfee_out`, na.rm = TRUE),
          sd_stats = sd(`2023_24_tuitionfee_out`, na.rm = TRUE),
          min_stats = min(`2023_24_tuitionfee_out`, na.rm = TRUE),
          max_stats = max(`2023_24_tuitionfee_out`, na.rm = TRUE)
              )
```

-> Why I calculated this and its importance at the bottom after all the calculations are done!

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2022_23_undergrad_enrollment`, na.rm = TRUE),
          sd_stats = sd(`2022_23_undergrad_enrollment`, na.rm = TRUE),
          min_stats = min(`2022_23_undergrad_enrollment`, na.rm = TRUE),
          max_stats = max(`2022_23_undergrad_enrollment`, na.rm = TRUE)
              )
```

-> Why I calculated this and its importance at the bottom after all the calculations are done!

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2022_23_tuitionfee_in`, na.rm = TRUE),
          sd_stats = sd(`2022_23_tuitionfee_in`, na.rm = TRUE),
          min_stats = min(`2022_23_tuitionfee_in`, na.rm = TRUE),
          max_stats = max(`2022_23_tuitionfee_in`, na.rm = TRUE)
              )
```
-> Why I calculated this and its importance at the bottom after all the calculations are done!

```{r}
tuition_statistics |>
  summarize(mean_stats = mean(`2022_23_tuitionfee_out`, na.rm = TRUE),
          sd_stats = sd(`2022_23_tuitionfee_out`, na.rm = TRUE),
          min_stats = min(`2022_23_tuitionfee_out`, na.rm = TRUE),
          max_stats = max(`2022_23_tuitionfee_out`, na.rm = TRUE)
              )
```
What did I just calculate, and why is it important?

ANSWER: I calculated the mean, standard deviation, minimum and maximum of undergraduate enrollment statistics from the 2023-24 school year, as well as in and out of state tuition fees. I did the exact same thing for the 2022-23 school year as well. This is going to be important because I am going to use all of this information to help determine my hypothesis of whether or not out of state tuition fees directly affect enrollment at universities. Finding the means of this data will help me get a gauge of all the schools clearly. 

Newsworthy Hypothesis Proposal:

ANSWER: Lower out of state tuition fees lead to higher enrollment totals, thus giving more enrollment money after. 

null hypothesis: The mean undergraduate enrollment is the same when comparing colleges with below the median out-of-state tuition vs. colleges that are above the median out-of-state tuition. 

Alternative hypothesis: The mean undergraduate enrollment increases when schools have out-of-state tuition lower than the media, rather than school with out-of-state tuition higher than the median. 


Questions that I might have:

ANSWER: The two major questions that I do have are if there is data that goes beyond 2021-22? Also, how should I really go about data that is not available in the graph? All in all, as an out of state student at UMD, granted that I was lucky enough to be granted a merit scholarship, I wonder how many students that had the academics to come to UMD couldn't because of out-of-state tuition prices being so high. How is enrollment decided? How does out of state compared to in state? And does it have a direct connection to enrollment? I would also really like to know what percentage of the total enrollment at these schools are in state, vs the ones that are out of state. 

Test Number One: One Sample t-test

```{r}
overall_mean <- tuition_statistics |>
  summarize(mean_stats = mean(`2023_24_undergrad_enrollment`, na.rm = TRUE)
  )
```

```{r}
alpha_level <- 0.05 


t_test_result <- t.test(
  tuition_statistics$`2023_24_undergrad_enrollment`,
  mu = 2980,
  alternative = "greater"
)

t_test_result
```

Test Number Two: Correlation test with a Scatter Plot

```{r}
tuition_statistics |> summarize(correlation = cor(`2023_24_tuitionfee_out`,`2023_24_undergrad_enrollment`, method="pearson", use = "complete.obs"))
```

```{r}

tuition_statistics |> 
  ggplot() +
  geom_point(aes(x=`2023_24_tuitionfee_out`, y=`2023_24_undergrad_enrollment`)) +
  geom_smooth(aes(x=`2023_24_tuitionfee_out`, y=`2023_24_undergrad_enrollment`), method="lm") +
  expand_limits(x = 0, y = 0)

```

Explanation of the data: 

For starters, I decided that my first test for the second part of the project was going to be a t-test. I wanted it to be a t-test so I can accurately determine if there is a relationship for what I was trying to prove in my hypothesis, which is that schools with a higher out of state tuition have less enrollment. A t test was going to help me prove for or against my hypothesis because a t-test helps to prove whether or not there was actually any significance between out-of-state tuition and enrollment in the 2023-24 school year. Then, getting a p-value was going to help me see if the answer I got just happened to be by random chance. The second test that I completed was a correlation test. I wanted to do a correlation test because it was going to tell me the linear relationship between the two variables I am testing: out-of-state tuition and enrollment. Now, let's look at the data starting with the correlation test. After, completing the test, I got a correlation of 0.08. This immediately to me pointed out that there was NOT a linear relationship between the two variables I was testing in my hypothesis. Next, looking at my t-test, I obtained a p-value of 0.499. This also immediately indicated to me that there was zero statistical significance at all the there is a true relationship between out-of-state tuition and enrollment, proving my alternate hypothesis to be wrong. 
