---
title: "Final Project - Part 1"
output: html_notebook
---

```{r}
library(tidyverse)
```
```{r}
tuition_statistics <- read_csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv") 
```
Direct link to the data descriptions:https://cran.r-project.org/web/packages/collegeScorecard/collegeScorecard.pdf

Direct source link: https://collegescorecard.ed.gov/

What does it contain and what time period does it cover?

ANSWER: The data contains 255 different universities, providing various years of data throughout. Each university is provided with data such as the state in which the university resides. Furthermore, each data set for the universities provide the in-state tuition fee, out-of-state tuition fee, and the total undergraduate enrollment for the school years 2021-2022, 2022-2023, and 2023-2024. 

Which columns will be most important for your analysis?

ANSWER: The columns that are going to be most important toward my analysis are honestly all of them. I am going to be deep diving into how the differences of in-state and out-of-state tuition fees affect the total enrollment of different universities. So, all of the data columns that have been provided are going to be an important factor in terms of my analysis. The only one that I probably won't dive too deep are the states in which the universities reside. 

What's missing that you might need or want?

ANSWER: Honestly, something that is missing that I might want to include is data about the specific in and out of state tuition enrollments. More specifically, I would really like to have data about who in terms of the total enrollment are in-state, vs who is out-of-state. 

